---
title: 'Module # 12 R Markdown'
author: "Loan Nguyen"
date: "3/21/23"
output: html_document
---

---

### ***An instruction***
Create a time series using gglpot2 and discuss the findings

---

### ***Load the libraries***
Add the necessary libraries

```{r}
library(DAAG)
library(stringr)
library(reshape2)
library(ggplot2)
```

### ***Load the data***
Using any dataset from any website or package or creating your own

```{r}
data(greatLakes)
data <- greatLakes
```

### ***Data munging***
Clean the data however you want, or don't clean it if it isn't necessary

```{r}
# rename columns
colnames(data)
colnames(data) <- str_to_title(colnames(data))
colnames(data) <- paste("Lake", colnames(data))
colnames(data)[4] <- "Lake St. Clair"
colnames(data)[2] <- "Lake Michigan-Huron"

# Create a data frame containing "Year" and data
data <- data.frame(Year = c(1918:2009), data)

# Transform the data in long format using the specified column name as the id parameter
melted <- melt(data, id = "Year")
# view the first few rows of melted
head(melted)

# Find the row numbers that meet the condition
rownames(melted)[which(melted$variable == "Lake.Ontario",)]
# Remove the rows
melted2 <- melted[-c(185:276),]
# Keeping rows
melted3 <- melted[c(185:276),]
```

### ***Create a plot***
Graph time series

```{r}
a <- ggplot(melted2, aes(Year, value, color = variable)) + geom_line() + ylab("Average Depth (m)") + geom_smooth(method = "lm", size = .5, linetype = "dashed") + annotate("text", x = 2000, y = 177.5, label = "dashed line = lm")
b <- ggplot(melted3, aes(Year, value, color = variable)) + geom_line() + ylab("Average Depth (m)") + scale_color_manual(values = "purple") + geom_smooth(method = "lm", size = .5, linetype = "dashed")
library(patchwork)
a + theme(legend.position = "bottom", legend.key.size = unit(.7, 'cm')) +
  guides(color = guide_legend(nrow = 2, byrow = T)) +
  b + theme(legend.position = "bottom", legend.key.size = unit(.7, 'cm')) +
  plot_annotation("Average Depth of Lakes", theme = theme(plot.title = element_text(hjust = 0.5)))
```

### ***Analysis***
Review the graph and discuss your findings: Do you notice any patterns?

> It’s shocking that Lake Ontario’s water level is low compared to the other three lakes. With the regression lines added, I can see trends by looking at their slope. The slope of Lake Michigan-Huron is slightly positive, while the slope of the other two lakes has the same slope, so I can safely assume that their average water levels in the future will gradually increase. It is evident that Lake Ontario has a very positive slope; this is bad because I can say with certainty that the average depth of Lake Ontario will increase rapidly, like two or three times compared to its normal level, in the future.
